### 2018-11-12 presentation/Summer Report/main.tex

\documentclass{beamer}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\usetheme{Cuerna}
\usecolortheme{default}

\title{Summer Work - A Brief Report}
\author{Shayan Amani\\}

\date{Summer 2018}
\institute{Department of Computer Science, University of New Hampshire}

\begin{document}

  \begin{frame}
    \titlepage
  \end{frame}

\begin{frame}
    \frametitle{Pneumonoultramicroscopicsilicovolcanoconiosis}
    \framesubtitle{Goals}
    Discussed during weekly meetings to allocate time to:
    \begin{itemize}
    \item familiarity with Species Distribution Models (SDM)
    \item examining Convexity of those models
    \item learning tools such as Convex Optimization
  \end{itemize}
  \end{frame}

  \begin{frame}
  	\frametitle{Accomplishments}
  \end{frame}
  \begin{itemize}
    \item finding out the methods like MaxLike and MaxEnt and some hands on experience with them
    \item 
  \end{itemize}
  \begin{frame}
  	\frametitle{Rec. Activities}
    Daily:
    \begin{itemize}
    \item push-ups x20
    \item 8 min abs workout
    \item summer fruits
    \end{itemize}
    Occasionally:
    \begin{itemize}
    \item White Mountains group hiking
    \item New Castle's beach on 4th of July
    \end{itemize}
	\includegraphics[scale=0.45]{img.jpg}
  \end{frame}

\end{document}