### 2018-11-13 presentation/PAC Optimal MDP Planning/main.tex
Pareto and Dirichlet distributions added to minutia

% Golden Spiral can be approximated with Golden ratio or Fibonacci sequence.
% Golden Ratio: a+b/a = a/b

\documentclass{beamer}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

\usetheme{Cuerna}
\usecolortheme{bluesimplex}
% default, bluesimplex, lettuce, brick

\title{PAC Optimal MDP Planning}
\author{Shayan Amani}

\date{fall 2018}
\institute{Department of Computer Science, University of New Hampshire}

\begin{document}

  \begin{frame}
    \titlepage
  \end{frame}

\begin{frame}
    \frametitle{Pneumonoultramicroscopicsilicovolcanoconiosis}
    \framesubtitle{SUB}

  \end{frame}

	%\includegraphics[scale=0.45]{img.jpg}

\end{document}