### 2018-11-19 AML-RL/FinalProject/AMLproject.tex
Updates from ShareLaTeX

\documentclass[a4paper,12pt]{article}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}

\usepackage{graphicx}
\usepackage{csquotes}
\usepackage[style=verbose-ibid,backend=bibtex]{biblatex}
\bibliography{sample}

\title{Advanced Machine Learning - Final Project}

\author{Shayan Amani}

\begin{document}
\maketitle

\section{Questions}
\begin{itemize}
    \item 2000 samples for 10 times of 200 years. What kind of times?
    \item 
\end{itemize}

\section{Definition}
We have 2000 samples for 200 years.

\section{Methodology}
We can use either On-Policy or Off-Policy methods.
\end{document}