### 2018-12-03 AML-RL/FinalProject/AMLproject.tex

\documentclass[a4paper,12pt]{article}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}

\usepackage{graphicx}
\usepackage{csquotes}
\usepackage[style=verbose-ibid,backend=bibtex]{biblatex}
\bibliography{sample}

\title{Advanced Machine Learning - Final Project}

\author{Shayan Amani}

\begin{document}
\maketitle

\section{Questions}
\begin{itemize}
    \item 2000 samples for 10 times of 200 years. What kind of times?
    \item What are states? How many?
\end{itemize}

\section{Definition}
We have 2000 samples for 200 years.

\subsection{State}
each row

\subsection{Action}
2 actions: 0 and 10.

\subsection{Reward}
A summation according to it's definition.

\section{Methodology}
We can use either On-Policy or Off-Policy methods. From the given samples it could be simply inferred that a model-free algorithm is needed and the reason is that the problem doesn't provide the transition probabilities.
\end{document}