### 2019-10-24 minutia/paper/paper.tex

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2019

% ready for submission
% \usepackage{neurips_2019}

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
    \usepackage[preprint]{neurips_2019}

% to compile a camera-ready version, add the [final] option, e.g.:
    %  \usepackage[final]{neurips_2019}

% to avoid loading the natbib package, add option nonatbib:
%     \usepackage[nonatbib]{neurips_2019}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

 \usepackage[sort]{natbib}

\title{Feature Extraction and Aggregation}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to LaTeX to determine where to break the
% lines. Using \AND forces a line break at that point. So, if LaTeX puts 3 of 4
% authors names on the first line, and the last on the second line, try using
% \AND instead of \And before the third author name.

\author{%
  Shayan Amani, Marek Petrik \\ %\thanks{Use footnote for providing further information
%    about author (webpage, alternative address)---\emph{not} for acknowledging
%    funding agencies.}%
  Department of Computer Science\\
  University of New Hampshire \\
  \texttt{\{shayan\}\{mpetrik\}@cs.unh.edu} \\
  % examples of more authors
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \AND
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
  % \And
  % Coauthor \\
  % Affiliation \\
  % Address \\
  % \texttt{email} \\
}

\begin{document}

\maketitle

\begin{abstract}
  Promising advances in deep reinforcement learning suggests several application to the algorithms. However robustness of these algorithms remains doubts to trust these algorithms in practice. In this paper we study resistance of the state-of-the-art algorithms to environment dynamics. We then show the amount of robustness we can achieve as a guarantee of performance in presence of those changes.
\end{abstract}

\section{Introduction}

Reassuring achievements in the Deep RL in a variety of fields absorb the attention from the community \citep{Mnih2015, Schulman2015TrustOptimization, Lillicrap2015ContinuousDDPG, silver2016a}. The cross fertilization in using approximation techniques by Deep Neural Networks in solving immense state space problems in RL is a key. While the newly emerged methods are growing at a staggering pace, a parallel area of research has been formed wherein a variety of techniques have been utilized to make these methods contravene their promised functionality. These efforts would be served to develop robustness in those spectacular models and reduce their susceptibility to data tampering or dynamics of the environment. Robustness ---as a property in machine learning--- elevates assurance about methods performance and their reliability to the foreground. In a few studies \cite{Goodfellow2014}  vulnerabilities of DNN's to adversarially-manipulated inputs are revealed. Therefore, robust algorithms, have to be of paramount importance in applying deep RL to real world scenarios to mitigate risks of implementation. Adversarial attacks as a way to perturbing these algorithms, are introduced in recent works \cite{Huang2017, Kos2017} from different aspects to degrade algorithms' performance. One of such potentials are changes in the environment dynamics.

In this work, we scrutinize state-of-the-art deep RL algorithms resiliency to environment parameters alterations and then introduced our approach to overcome those. Then a guarantee of performance to these attacks is introduced.

\section{Related Work}
\cite{Goodfellow2014} stated and consequently depicted that the root cause of vulnerability to adversarial examples laid behind the piecewise linearity attributes of neural network models.

\section{Definition}
For the sake of clarity, we elaborate on some preliminary definitions here to have a consistent set of notations to the end of this paper.

\subsection{Adversarial attacks}
The designed attack may have different levels of access to the underlying training policy which have motivated the community to classify adversarial attacks into two distinct categories, namely \textit{white-box} or \textit{black-box} attacks. White-box attacks illustrating the scenario in which the... Adversarial attacks fall in either one of the following categories based upon the stage which in the adversary perturbs the model:
\begin{itemize}
    \item Train time: Data poisoning.
    \item Evaluation time: Adversarial examples.
    \item Deploy time: Black-box attacks.
\end{itemize}

\subsection{Objective}
The ultimate goal is deceiving the model either in choosing the worst possible action in an RL task or misclassifying an input in a classification task in a way which is not perceptible for unaided eye. In terms of a Markovian RL task we show how minimum set of worst-case actions are needed to result in the minimum final reward. We try to show that the model is not robust to other configuration of each benchmarks with different dynamics.

The main goal of adversarial training and examples can be expressed briefly in form of the following equation:
\begin{equation}
\begin{aligned}
& \underset{\theta}{\min}
& & \sum_{x,y \in S} \underset{\delta \in \Delta}{\max}
& & & \textbb{Loss}(x + \delta, y; \theta)
\end{aligned}
\end{equation}

\section{Devising Adversarial Attacks}
Neural networks are vulnerable to adversarial attacks either in supervised learning or reinforcement learning settings \cite{Huang2017}. Such weakness could possibly result in unrecognizable differences for the unaided eye in the presence of perturbation which leads to misclassification for neural network based model \cite{Szegedy2014, Goodfellow2014}.

\subsection{Observation-perturbed attacks}

where in we trying to find an adversarial example (inner maximization) in order to adversarially train a robust model (outer minimization).

We define $s_\eta$ as the perturbed observation:
\begin{equation}
    s_\eta = s . (1 + \epsilon) 
\end{equation}

Let $s_\eta$ be the perturbed observation and $s_c$ scaled observation, so we are trying to find the following:
\begin{equation}
\begin{aligned}
& \underset{s}{\argmin}
& & {\norm{s_\eta - s_c}} = \norm{s . (1 + \eps) - s_c}
\end{aligned}
\end{equation}
which is the state observation that have the minimum distance between real-world [scaled] observation and the perturbed one.

In order to expand the applicability of designed adversarial attacks to an assortment of algorithms [models], we rely on \textit{transferability property} of RL-specific adversarial attacks \cite{Szegedy2014, Papernot2016, Goodfellow2014a}. \cite{Huang2017} clearly assert [or asserts?!] attacks are deployable to a group of trained RL models aimed at the same task.

\subsection{Reward-Perturbed attacks}
In an adversarial environment we try to perturb reward to interfere in learning process of the RL agent. We considered this attack also in scenarios with a single reward which we will show the process later.

\section{Adversarial attacks on different methodologies}
Here we review adversarial attacks on a variety of value function based state-of-art algorithms proposed by the community so far. As \cite{Pattanaik2018} pointed out, the concept of achieving a successful adversarial attack is unalike in reinforcement learning tasks compared to image classification tasks. Worst \textit{action} is tangible and feasible but not worst \textit{image}.

\subsection{Deep Q Learning (DQN)}
Utilizing experience replay and the leveraging target network, DQN take advantage of them to gain more stability \cite{Pattanaik2018}.

\section{Experiment}
We apply the approach to two MDPs proposed by \cite{Strehl2004AnProcesses}. \textit{RiverSwim} an MDP consists of six states and \textit{SixArms} one with seven states. In \textit{RiverSwim} flow of the river is to left and the swimmer picks probabilistically equal either state 1 or 2 as the start point. All states but two terminal states have zero reward to land at while the right most one rewards substantially higher than the other one. The agent in SixArms selects between six distinct actions which pulls different arms of a multi-armed bandit. By pulling an arm the agent traverses to a new state. Although the agent does not obtain reward by pulling the arms but going to the connected state to that arm is highly rewarding in which the transition probability is in inverse relationship with the reward value.

As both problems favor keeping the agent in states with lower reward, exploration is paramount to maximize the accumulated reward. This is also confirmed by the authors in \cite{Strehl2004AnProcesses} that the agent will fail to learn when a rudimentary $\epsilon$-greedy algorithm.

Our implementation uses Deep Q-Network (DQN) \cite{Mnih2015} as the policy optimizer. OpenAI Gym \cite{BrockmanOpenAIGym} is the testbed in the benchmarks in this paper. In training phase, models are trained in the environment. Online methods such as DQN learn by interacting directly in the environment, however, CRAAM

% \begin{figure}
%     \centering
%     \includegraphics[width=1\columnwidth]{minutia/cartpole-v1_batchsize-32_episodes-100_full-step.png}
%     \caption{Average reward of DQN training to 100 episodes in cartpole.}
%     \label{fig:my_label}
% \end{figure}

\section{Future Works}
We developed the presented approach with having reproducibility in mind. It could be served as a launching pad for the future contributors who want to surf this newly emerged field deeper. Future ideas to pursue is a generative model to be able to generate adversaries from a network.

\section*{Appendix}

\subsection{Implementation note}
* Reshape is needed after getting each state from Gym models.

\bibliographystyle{plain}
\bibliography{library}
% \section*{References}

% [1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms for
% connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and T.K.\ Leen
% (eds.), {\it Advances in Neural Information Processing Systems 7},
% pp.\ 609--616. Cambridge, MA: MIT Press.

% [2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS: Exploring
%   Realistic Neural Models with the GEneral NEural SImulation System.}  New York:
% TELOS/Springer--Verlag.

% [3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of learning and
% recall at excitatory recurrent synapses and cholinergic modulation in rat
% hippocampal region CA3. {\it Journal of Neuroscience} {\bf 15}(7):5249-5262.

\end{document}